PANCHANGE_analysis

### 2020-09-03 PANCHANGE_analysis.Rmd
Exclusions chunk moved again to just above models, so that data manipulation chunks only have to be run once.

# Psychiatric disorder hierarchy
------------
OLD VERSIONS:
control_numeric == 1 ~ 0,
depressive_disorders_numeric == 1 ~ 1,
anxiety_disorders_numeric == 1 ~ 2,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
eating_disorders_numeric == 1 ~ 4,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
mhd.personality_disorder_numeric == 1 ~ 7,
autism_spectrum_disorder_numeric == 1 ~ 8,
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
psychotic_disorders_numeric == 1 ~ 10

psychotic_disorders_numeric == 1 ~ 10
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
autism_spectrum_disorder_numeric == 1 ~ 8,
mhd.personality_disorder_numeric == 1 ~ 7,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
eating_disorders_numeric == 1 ~ 4,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
anxiety_disorders_numeric == 1 ~ 2,
depressive_disorders_numeric == 1 ~ 1,
control_numeric == 1 ~ 0,
------------

@@

## Hierarchy including psychosis
```{r Psychiatric disorders hierarchy including psychosis - dat.mhd}
dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical_psychosis_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 7,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical_psychosis =
      recode_factor(
        Disorder_hierarchical_psychosis_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders",
        `7` = "Psychotic disorders"
      )
  )

dat.raw %>%
  freq(Disorder_hierarchical_psychosis)
```

```{r Psychiatric disorders hierarchy without psychosis - dat.mhd}
dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical_numeric =
      case_when(
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat.raw <- dat.raw %>%
  mutate(
    Disorder_hierarchical =
      recode_factor(
        Disorder_hierarchical_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders"
      )
  )

dat.raw %>%
  freq(Disorder_hierarchical) 
dat.raw %>% 
  freq(Disorder_hierarchical_psychosis)
```

Investigate NAs 
NAs are coded for individuals that don't have one of the listed disorders in the hierarchy, but who have another diagnosis that excludes them from being a control (e.g. personality disorder, ADHD, ASD). These individuals will be excluded from the analyses.
```{r NAs in diagnosis}
dat.raw %>%
  filter(is.na(Disorder_hierarchical)) %>%
  freq(Sample)
```

```{r disorders total count}
dat.mhd %>%
  freq(disorders_total_count)
```

## Correlation matrix
Using dat.mhd for the correlation matrix.
```{r Correlation matrix of psychiatric disorders}
# Select items for correlation matrix
mhd.items <- c(
    "depressive_disorders",
    "anxiety_disorders",
    "eating_disorders",
    "obsessive_compulsive_disorders",
    "psychotic_disorders",
    "mhd.mania_hypomania_bipolar_or_manicdepression",
    "mhd.posttraumatic_stress_disorder_ptsd",
    "autism_spectrum_disorder",
    "mhd.attention_deficit_hyperactivity_disorder",
    "mhd.personality_disorder",
    "control"
)

# Calculate correlation matrix on specific columns (variables) from your data frame
mhd.corr.mat <- hetcor(
  as.data.frame(dat.mhd[,mhd.items]), # Specify the data
  use = "pairwise.complete.obs" # Use all pairwise observations
)

```{r gather to long format Disorder_hierarchical}
dat.long.psych <- dat.mhd %>%
  gather(
    key = "Disorder_inc_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    all_of(diff_score_cols),
    Disorder_inc_raw,
    Diagnosis
  ) %>%
  mutate(Disorder_inc =
           recode_factor(
             Disorder_inc_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             autism_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

@@

>>>>

# Psychiatric disorder hierarchy
- - - -
OLD VERSIONS:
control_numeric == 1 ~ 0,
depressive_disorders_numeric == 1 ~ 1,
anxiety_disorders_numeric == 1 ~ 2,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
eating_disorders_numeric == 1 ~ 4,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
mhd.personality_disorder_numeric == 1 ~ 7,
autism_spectrum_disorder_numeric == 1 ~ 8,
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
psychotic_disorders_numeric == 1 ~ 10

psychotic_disorders_numeric == 1 ~ 10
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
autism_spectrum_disorder_numeric == 1 ~ 8,
mhd.personality_disorder_numeric == 1 ~ 7,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
eating_disorders_numeric == 1 ~ 4,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
anxiety_disorders_numeric == 1 ~ 2,
depressive_disorders_numeric == 1 ~ 1,
control_numeric == 1 ~ 0,
- - - -

@@

```{r Psychiatric disorders hierarchy including psychosis}
dat <- dat.raw %>%
  mutate(
    Disorder_hierarchical_psychosis_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 7,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat <- dat.raw %>%
  mutate(
    Disorder_hierarchical_psychosis =
      recode_factor(
        Disorder_hierarchical_psychosis_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders",
        `7` = "Psychotic disorders"
      )
  )

dat %>%
  freq(Disorder_hierarchical_psychosis)
```

```{r Psychiatric disorders hierarchy without psychosis}
dat <- dat %>%
  mutate(
    Disorder_hierarchical_numeric =
      case_when(
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat <- dat %>%
  mutate(
    Disorder_hierarchical =
      recode_factor(
        Disorder_hierarchical_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders"
      )
  )

dat %>%
  freq(Disorder_hierarchical) 
dat %>% 
  freq(Disorder_hierarchical_psychosis)
```

Investigate NAs 
NAs are coded for individuals that don't have one of the listed disorders in the hierarchy, but who have another diagnosis that excludes them from being a control (e.g. personality disorder, ADHD, ASD). These individuals will be excluded from the analyses.
```{r NAs in diagnosis}
dat %>%
  filter(is.na(Disorder_hierarchical)) %>%
  freq(Sample)
```

@@

```{r disorders total count}
dat %>%
  freq(disorders_total_count)
```

## Correlation matrix
Using dat for the correlation matrix.
```{r Correlation matrix of psychiatric disorders}
# Select items for correlation matrix
mhd.items <- c(
    "depressive_disorders",
    "anxiety_disorders",
    "eating_disorders",
    "obsessive_compulsive_disorders",
    "psychotic_disorders",
    "mhd.mania_hypomania_bipolar_or_manicdepression",
    "mhd.posttraumatic_stress_disorder_ptsd",
    "autism_spectrum_disorder",
    "mhd.attention_deficit_hyperactivity_disorder",
    "mhd.personality_disorder",
    "control"
)

# Calculate correlation matrix on specific columns (variables) from your data frame
mhd.corr.mat <- hetcor(
  as.data.frame(dat[,mhd.items]), # Specify the data
  use = "pairwise.complete.obs" # Use all pairwise observations
)

```{r gather to long format Disorder_hierarchical}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_inc_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    all_of(diff_score_cols),
    Disorder_inc_raw,
    Diagnosis
  ) %>%
  mutate(Disorder_inc =
           recode_factor(
             Disorder_inc_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             autism_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))
